Ethnicity sensitive author disambiguation using semi-supervised learning

### 2016-01-12 paper/author_disambiguation.tex

\usepackage{xspace}
\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}

@@

\numberofauthors{4}
\author{
\alignauthor
Gilles Louppe\\
       \affaddr{CERN, Switzerland}\\
       \email{g.louppe@cern.ch}
\alignauthor
Hussein T. Al-Natsheh\\
       \affaddr{ISH-LYON (USR 3385), CNRS, France}\\
       \email{hussein.al-natsheh@ish-lyon.cnrs.fr}
\alignauthor
Mateusz Susik\\
       \affaddr{University of Warsaw}\\
       \email{msusik@student.uw.edu.pl}
\alignauthor
\and
Eamonn Maguire\\
       \affaddr{CERN, Switzerland}\\
       \email{eamonn.james.maguire@cern.ch}
}

@@

\section{Semi-supervised author disambiguation}
\label{methods}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig-pub-to-signature}
\caption{An example signature $s$ for "Doe, John". A \textit{signature} is
defined as unique piece of information identifying an author on a publication,
along with any other metadata that can be derived from it, such as publication
title, co-authors or date of publication.}
\label{fig:signature}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig-workflow}
\caption{Pipeline for author disambiguation: (a)
signatures are \textit{blocked} to reduce computational complexity, (b) a linkage
function is built with supervised learning, (c) independently within each block, signatures
are grouped using hierarchical agglomerative clustering.}
\label{fig:workflow}
\end{figure*}

@@

To account for these issues we propose instead to block signatures based on the
phonetic representation of the normalized surname.
Normalization involves stripping accents (\eg, "Jabłoński, Ł" $\rightarrow$ "Jablonski, L") and name
affixes that inconsistently appear in signatures (\eg, "van der Waals, J. D."
$\rightarrow$ "Waals, J. D."), while phonetization is based either on the
Double Metaphone \cite{doublemetaphone}, the NYSIIS \cite{nysiis} or the
Soundex \cite{Soundex} phonetic algorithms for mapping author names to their pronunciations.
Together, these processing steps allow for grouping of most name variants of the same
person in the same block with a small increase in the overall computational complexity, thereby solving case 1.

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig-cuts}
\caption{Semi-supervised cut-off strategies to form flat clusters of signatures.}
\label{fig:cuts}
\end{figure*}

>>>>

\usepackage{xspace}
\usepackage[hidelinks]{hyperref}
\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}

@@

\numberofauthors{4}
\author{
\alignauthor
Gilles Louppe\\
       \affaddr{CERN, Switzerland}\\
       \email{g.louppe@cern.ch}
\alignauthor
Hussein T. Al-Natsheh\\
       \affaddr{ISH-LYON (USR 3385), CNRS, France}\\
       \email{hussein.al-natsheh@ish-lyon.cnrs.fr}
\and
\alignauthor
Mateusz Susik\\
       \affaddr{University of Warsaw}\\
       \email{msusik@student.uw.edu.pl}
\alignauthor
Eamonn Maguire\\
       \affaddr{CERN, Switzerland}\\
       \email{eamonn.james.maguire@cern.ch}
}

@@

\section{Semi-supervised\\ author disambiguation}
\label{methods}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig-pub-to-signature.pdf}
\caption{An example signature $s$ for "Doe, John". A \textit{signature} is
defined as unique piece of information identifying an author on a publication,
along with any other metadata that can be derived from it, such as publication
title, co-authors or date of publication.}
\label{fig:signature}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig-workflow.pdf}
\caption{Pipeline for author disambiguation: (a)
signatures are \textit{blocked} to reduce computational complexity, (b) a linkage
function is built with supervised learning, (c) independently within each block, signatures
are grouped using hierarchical agglomerative clustering.}
\label{fig:workflow}
\end{figure*}

@@

To account for these issues we propose instead to block signatures based on the
phonetic representation of the normalized surname.
Normalization involves stripping accents (e.g., "Jab\l{}o\'nski, \L{}" $\rightarrow$ "Jablonski, L") and name
affixes that inconsistently appear in signatures (e.g., "van der Waals, J. D."
$\rightarrow$ "Waals, J. D."), while phonetization is based either on the
Double Metaphone \cite{doublemetaphone}, the NYSIIS \cite{nysiis} or the
Soundex \cite{Soundex} phonetic algorithms for mapping author names to their pronunciations.
Together, these processing steps allow for grouping of most name variants of the same
person in the same block with a small increase in the overall computational complexity, thereby solving case 1.

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{fig-cuts.pdf}
\caption{Semi-supervised cut-off strategies to form flat clusters of signatures.}
\label{fig:cuts}
\end{figure*}
